\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-08-31 notes/2loopNLSM.tex
More 2-loop NLSM and figures.

The second key constraint in an ansatz construction of color-dual numerators comes from graph symmetries.
Every numerator is required to respect all of the symmetries (or automorphisms) of its associated graph.
While it is very convenient to impose graph symmetries, it is not completely necessary in order to perform the double copy.
At tree level, finding manifestly color-dual numerators is often achieved at the expense of manifest Bose symmetry (see YZ theory, j-theory, Alex numerators from CHY, Henrik numerators).
Nonetheless, for a pure ansatz construction it is very reasonable to impose graph symmetries.
Note that the graph symmetry constraint is imposed on \emph{every} graph, not just the basis graphs.
For example, the crossed box,
\be
\CrossedBoxGraphSym ,
\ee
is unchanged upon reflection across the horizontal axis (shown as a red dashed line).
Since the crossed box can be written in terms of the double box and penta-triangle, this imposes non-trivial constraints on the ansatz.

@@

As we have seen, it is possible to coerce tree numerators into 1-loop numerators at any multiplicity for pions.
However, these methods cannot be reapplied to generate 2-loop numerators, which begs the question ``is it possible to find globally color-dual numerators beyond 1-loop''.
We will tackle this question for NLSM before moving on to YM.

Every vertex in NLSM scales as $k^2$ so the numerator of a 4pt 2-loop diagram diagram scales as $k^{12}$.
At this mass dimension there are 8,008 independent dot products of momenta.
Ignoring tadpoles and BEL graphs, there are 14 cubic 4pt 2-loop graphs, but they are related to one another via 21 distinct Jacobi relations.
As mentioned in (review section), this means that the numerator of every graph can be expressed in terms of the double box and penta-triangle.
With two basis graphs there is a total of 16,016 ansatz parameters.\footnote{\texttt{Mathematica} is not suited to equations with sixteen thousand parameters so a different method is desirable.
One possibility is to use a solver designed for integral reduction.
However, these solvers typically assume that the matrix has a very low density which is not the case for equations coming from NLSM which typically have a density of a few percent.
While the equations in YM naturally separate into different ``helicity'' sectors (or $e_i \cdot e_j$ and $e_i \cdot k_j$ sectors in $D$ dimensions), there is no such separation of sectors in a scalar theory where any product of momenta can interfere with any other.
The NLSM constraint equations were solved using a custom solver, \texttt{FiniteFieldSolve}, designed to solve large linear systems of arbitrary density.
\texttt{FiniteFieldSolve} will be released publicly shortly.}
Imposing all of the Jacobi relations eliminates all but 4,473 of the original free parameters.
Further imposing graph symmetries reduces the number of free parameters to 1,243.

The final essential constraint is unitarity.
Cuts (b) and (c) of \cref{eq:PotentialCuts} are the only non-vanishing cuts that need to be enforced.  \jm{Move some of the cut pictures here?}
This means that even at two loops, unitarity cuts are still only probing the on-shell 4pt NLSM amplitude.
Even though (b) and (c) are the only non-zero cuts, other cuts, for example the maximal cut (d), must also be imposed in order to correctly reproduce NLSM.
The maximal cut vanishes because only 3pt amplitudes contribute and these amplitudes vanish in any scalar theory.
%It is possible to automatically enforce (or ``bake in'') the double box and penta-triangle max cuts on the ansatz by writing the ansatz in terms of inverse propagators (and remaining irreducible scalar products) and then only keeping terms that involve at least one inverse propagator since those terms will vanish on those max cuts, just as the product of 3pt scalar amplitudes will.
%However, for the sake of simplicity we do not take this route.
Imposing all relevant cuts leaves 765 free parameters representing the generalized gauge freedom in the answer.

@@

At this point color-kinematics duality has been satisfied, so the double copy will proceed without issue.
However, the plethora of generalized gauge freedom parameters leaves the option for enforcing additional aesthetic constraints.
$\mathcal{N}=4$ SYM provides several hints for further conditions to impose on the ansatz.
For example, for maximally supersymmetric gauge theory it is possible to enforce the no triangle hypothesis, manifest loop power counting, and, at least for 4pt and up to four loops, strip off a factor of $st A_\text{tree}$ from the integrand (I'm not sure I know all of the papers to cite here).
For the NLSM integrand it is desirable to make color-kinematics duality as manifest as possible.
One hope would be to factor out some piece of the 1-loop numerator since this at least manifests antisymmetry for the vertices involving external legs.
However, a more fruitful direction is to match onto the one known theory with pion power counting that manifests color-kinematics duality to all loop order, namely, Zakharov-Mikhailov theory (cite ZM, James).

@@

The loop momentum structure of ZM theory provides one final hint for simplifying the pion numerators.
Both loop momenta $\ell_1$ and $\ell_2$ appear in the ZM numerators but only 5 to 8 powers appear in any term, that is, each term in the ZM numerator scales as either $\ell^5$, $\ell^6$, $\ell^7$, or $\ell^8$ where $\ell$ here refers to any combination of $\ell_1$ or $\ell_2$.
When the pion numerator is also forced to have this loop power counting, the number of generalized gauge freedom parameters reduces to 58.
The pion numerator appearing in the ancillary files contains only these 58 degrees of freedom.\footnote{When the 4pt pion amplitude is normalized to $-k_1\cdot k_2$ and the ZM vertex is normalized to $k_1^\mu \epsilon_{\mu\nu}k_2^\nu$, the parameter $z$ is fixed to $216/565$, which is clearly unrelated to the difference in coupling constants.}

>>>>

The second key constraint in an ansatz construction of color-dual numerators comes from graph symmetries.
Every numerator is required to respect all of the symmetries (or automorphisms) of its associated graph.
While it is very convenient to impose graph symmetries, it is not completely necessary in order to perform the double copy.
At tree level, finding manifestly color-dual numerators is often achieved at the expense of manifest Bose symmetry (see YZ theory, j-theory, Alex numerators from CHY, Henrik numerators).
Nonetheless, for a pure ansatz construction it is very reasonable to impose graph symmetries.
Note that the graph symmetry constraint is imposed on \emph{every} graph, not just the basis graphs.
%For example, the symmetries of the crossed box are spanned by
%\be
%\CrossedBoxGraphSym{1}{2}{3}{4} \sim \CrossedBoxGraphSym{2}{1}{3}{4} \text{ and } \CrossedBoxGraphSym{1}{2}{3}{4} \sim \CrossedBoxGraphSym{2}{1}{4}{3}
%\ee
For example, the symmetries of the crossed box
\be
\CrossedBoxGraphSym{1}{2}{3}{4}
\ee
are spanned by $\{1234\} \to \{2134\}$ and $\{1234\} \to \{1243\}$.
In general, internal edge labels will have to be tracked as well.
Care must be taken with the signs in a symmetry relation.
To compensate for the antisymmetric structure of $f^{abc}$, the vertices are all totally antisymmetric and hence have signs built into their orientations.
Once the signs are included in the symmetry relations, the crossed box can be expressed in terms of the double box and penta-triangle, thus imposing non-trivial constraints on the ansatz.

@@

As we have seen, it is possible to coerce tree numerators into 1-loop numerators at any multiplicity for pions.
However, these methods cannot be reapplied to generate higher-loop numerators, which begs the question ``is it possible to find globally color-dual numerators beyond 1-loop''.
We will tackle this question for N$^2$LO 2-to-2 scattering in NLSM before moving on to the same process in YM.

Ignoring tadpoles and BEL graphs, there are 14 cubic 4pt 2-loop graphs (see later picture) which are related to each other via 21 Jacobi relations.
As mentioned in (review section), color-kinematics duality ensures that the numerator of every graph can be expressed in terms of a basis of the double box and penta-triangle,
\be 
\NLSMBasisDoubleBox \text{ and } \NLSMBasisPentaTriangle .
\ee
For a scalar theory like NLSM the numerator only depends on dot products of momenta.
The basis of Mandelstam invariants for both of these graphs can be taken to be
\be
\{ k_{13}, k_{23}, k_{15}, k_{16}, k_{25}, k_{26}, k_{35}, k_{36}, k_{55}, k_{56}, k_{66} \},
\ee
where $k_{ij}$ means $k_i \cdot k_j$, $k_5=\ell_1$, and $k_5 = \ell_2$.
%%%%
% Basis for both diagrams
% {d[p[1], p[3]], d[p[1], p[6]], d[p[1], p[7]], d[p[2], p[3]],  d[p[2], p[6]], d[p[2], p[7]], d[p[3], p[6]], d[p[3], p[7]],  d[p[6], p[6]], d[p[6], p[7]], d[p[7], p[7]]}
%%%%
Every vertex in NLSM scales as $k^2$ so the numerator of each diagram scales as $k^{12}$.
At this mass dimension there are 8,008 independent dot products of momenta.
With two basis graphs there is a total of 16,016 ansatz parameters.\footnote{\texttt{Mathematica} is not suited to equations with sixteen thousand parameters so a different method is desirable.
One possibility is to use a solver designed for integral reduction.
However, these solvers typically assume that the matrix has a very low density which is not the case for equations coming from NLSM which typically have a density of a few percent.
While the equations in YM naturally separate into different ``helicity'' sectors (or $e_i \cdot e_j$ and $e_i \cdot k_j$ sectors in $D$ dimensions), there is no such separation of sectors in a scalar theory where any product of momenta can interfere with any other.
The NLSM constraint equations were solved using a custom solver, \texttt{FiniteFieldSolve}, designed to solve large linear systems of arbitrary density.
\texttt{FiniteFieldSolve} will be released publicly shortly.}
Imposing all of the Jacobi relations eliminates all but 4,473 of the original free parameters.
Further imposing graph symmetries reduces the number of free parameters to 1,243.

The final essential constraint is unitarity.
Cuts (b) and (c) of \cref{eq:PotentialCuts} are the only non-vanishing cuts that need to be enforced.  \jm{Move some of the cut pictures here?}
This means that even at two loops, unitarity cuts are still only probing the on-shell 4pt NLSM amplitude.
Even though (b) and (c) are the only non-zero cuts, other cuts, for example the maximal cut (d), must also be imposed in order to correctly reproduce NLSM.
The maximal cut vanishes because only 3pt amplitudes contribute and these amplitudes vanish in any scalar theory.
%It is possible to automatically enforce (or ``bake in'') the double box and penta-triangle max cuts on the ansatz by writing the ansatz in terms of inverse propagators (and remaining irreducible scalar products) and then only keeping terms that involve at least one inverse propagator since those terms will vanish on those max cuts, just as the product of 3pt scalar amplitudes will.
%However, for the sake of simplicity we do not take this route.
Imposing all relevant cuts leaves 765 free parameters representing the generalized gauge freedom in the answer.
Again, ambiguous non-physical cuts are not imposed.

@@

At this point color-kinematics duality has been satisfied, so the double copy will proceed without issue.
However, the plethora of generalized gauge freedom parameters leaves the option for enforcing additional aesthetic constraints.
$\mathcal{N}=4$ SYM provides several hints for further conditions to impose on the ansatz.
For example, for maximally supersymmetric gauge theory it is possible to enforce the no triangle hypothesis, manifest loop power counting, and, at least for 4pt and up to four loops, strip off a factor of $st A_\text{tree}$ from the integrand (I'm not sure I know all of the papers to cite here).
For the NLSM integrand it is desirable to make color-kinematics duality as manifest as possible.
One hope would be to factor out some piece of the 1-loop numerator since this at least manifests antisymmetry for the vertices involving external legs.
However, a more fruitful direction is to match onto the one known theory with pion power counting that manifests color-kinematics duality to all loop order, namely, Zakharov-Mikhailov theory \citepls{Original ZM paper, ZM CK paper}.

@@

The loop momentum structure of ZM theory provides one final hint for simplifying the pion numerators.
A generic term in the ZM numerator (of either basis graph) looks schematically like
\be
\ell_1^m \ell_2^n k^{12-(m+n)} \text{ where } 5 \leq m+n \leq 8.
\ee
In other words, every term contains between 5 and 8 powers of loop momenta, as one would expect from a set of local, cubic Feynman rules.
\jm{I believe this is true but I need to check it.}
When the pion numerator is also forced to have this loop power counting, the number of generalized gauge freedom parameters reduces to 58.
The pion numerator appearing in the ancillary files contains only these 58 degrees of freedom.\footnote{When the 4pt pion amplitude is normalized to $-k_1\cdot k_2$ and the ZM vertex is normalized to $k_1^\mu \epsilon_{\mu\nu}k_2^\nu$, the parameter $z$ is fixed to $216/565$, which is clearly unrelated to the difference in coupling constants.}
